A Graph Theoretic Approach to Problem Formulation for Multidisciplinary Design Analysis and Optimization

### 2013-08-07 mcg_fpg.tex
files for submitted version and reviewer 3's comments

\section{The MCG and FPG}
    \label{s:building graphs}
We now discuss how the graph-based syntax defined in Sec.~\ref{s:syntax definition} is used to represent an MDAO problem in its entirety. This discussion is based on the MCG and the FPG.

\subsection{Maximal Connectivity Graph}
    \label{ss:MCG}
For a specified MDAO problem, the maximal connectivity graph represents every analysis tool, objective, constraint, and input being considered, as well as every possible interconnection among them.
The definition of the MCG is given via construction.
    To construct the maximal connectivity graph, we presume that a set of analysis tools, inputs, objectives, and constraints are provided. 
    Each of the $m$ analysis codes correspond to an index $i\in I_A$, $I_A=\{1,2,\ldots,m\}$, and are represented by an analysis block graph $G_{A_i}=(V_{A_i},E_{A_i})$.
    Each of the $n$ expression blocks correspond to an index $i\in I_E$, $I_E=\{1,2,\ldots,n\}$, and are represented by an expression block graph $G_{E_i}=(V_{E_i},E_{E_i})$.
    Finally, the inputs are represented as a set of variable nodes $V_\txt{in}$.
    We presume that $V_\txt{in}$, each $G_{A,i}$, and each $G_{E,i}$ are given, and that any potential connection between variables is specified in the form of connection edges in the set $E_{M,\txt{con}}$. One method for defining these connections is to use a consistant variable naming convention.
    We may then construct the maximal connectivity graph $M=(V_M,E_M)$ as
    \begin{IEEEeqnarray*}{rCl}
    V_M & = & V_\txt{in} \cup \left( \bigcup_{i \in I_A} V_{A_i} \right) \cup \left( \bigcup_{i \in I_E} V_{E_i} \right), \\
    E_M & = & E_{M,\txt{con}} \cup \left( \bigcup_{i \in I_A} E_{A_i} \right)  \cup \left( \bigcup_{i \in I_E} E_{E_i} \right).
    \end{IEEEeqnarray*}
    The MCG $M$ is uniquely determined by the given set of analysis blocks, inputs, and expression blocks. 
    In the cases where the set of inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.

The nodes and edges in $M$ can be partioned in sets according to their type:
\begin{equation}
V_M = V_{M,\txt{var}} \cup V_{M,\txt{mod}}, \ V_{M,\txt{var}} \cap V_{M,\txt{mod}} = \emptyset,
\end{equation}
where $V_{M,\txt{var}}$ and $V_{M,\txt{mod}}$ are the sets of variable nodes and function nodes, respecively;
\begin{equation}
E_M = E_{M,\txt{con}} \cup E_{M,\txt{mod}}, \ E_{M,\txt{con}} \cap E_{M,\txt{mod}} = \emptyset,
\end{equation}
where $ E_{M,\txt{con}}$ and $E_{M,\txt{mod}}$ are the sets of connection edges and function edges, respectively. These sets will be referenced in the process for obtaining an FPG.

\subsection{Fundamental Problem Graph}
    \label{ss:FPG}
    We now define the fundamental problem graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions:
    \begin{enumerate}
    \item[(1)] $F \subset M$ and $G_{E_i} \subset F  \ \forall i \in I_E$
%    \item[(1)] $F \subset M$ and $V_\txt{out} \subset V_F$
    \item[(2)] $\displaystyle{\forall i \in I_A, \txt{ if } F \cap G_{A_i} \neq \emptyset \txt{ then } G_{A_i} \subset F}$
%    \item[(2)] $\displaystyle{\forall i \in I, \txt{ if } F \cap A_i \neq \emptyset \txt{ then } A_i \subset F}$
    \item[(3)] $\displaystyle{\forall v \in V_F \txt{ with }v \in V_{M,\txt{var}} \txt{ and }  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_{E_i}}$
%    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_\txt{out}}$
    \end{enumerate}
    %The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). 
    %Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
    %Requirement (3) stipulates that only the inputs that are being used should be included. 
    %Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
    Condition (1) asserts that only the nodes and edges provided by the maximal connectivity graph can be used in the fundamental problem graph and that every expression block must be included. 
    Condition (2) requires that analysis blocks must be included or excluded in their entirety. 
    Condition (3) stipulates that the number of edges directed into each variable node must be within the lower and upper in-degree limits; 
    if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is the location of a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is the location of a \emph{collision}.
    Lastly, condition (4) ensures that only the nodes that are being used are included in the FPG by requiring that for every node a reverse path exists from at least one expression block to that node. 
    The reverse graph $R_F$ is obtained from $F$ by simply switching the orientation of every edge in $E_F$.

@@

\subsection{Algorithm for obtaining the Fundamental Problem Graph from the MCG}
    \label{ss:obtaining FPG}
    In general, there may be zero, one, or many graphs that satisfy the FPG conditions in Sec.~\ref{ss:FPG}.
    Here, we describe an algorithm for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. 
    With this approach, the problem is reduced to deciding which connection edges to remove.

@@

        To begin the process, an initial graph is created as a copy of the MCG
        \begin{equation}
        F_0 = M,
        \end{equation}
        with $F_0 = (V_{F_0},E_{F_0})$, and $C_{F_0,\txt{con}}$ is the set of connection edges. 
        The set of variable nodes that are holes is identified as
        \begin{equation}
        H = \{v \in V_{F_0} \st v \in V_{M,\txt{var}} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
        \end{equation}
        which is the set of variable nodes with fewer incoming edges than are allowed by the lower indegree limit.
        The updated set of connection edges is then created by removing the edges preceding or succeeding the analysis block:
        \begin{equation}
        E_{F_1,\txt{con}} = E_{F_0,\txt{con}} \setminus \{(x,y) \in E_{F_0,\txt{con}} \st x \in V_{A_i} \txt{ or } y \in V_{A_i}, \txt{ and } H\cap V_{A_i} \neq \emptyset \}.
    \end{equation}
Because removing these edges can create new holes, this step must be repeated until no additional holes are found.
        If the hole identification step identifies a variable node in an expression block as hole, meaning $V_{E_i} \cap H \neq \emptyset$ for some $i \in I_E$, then the algorithm terminates because an FPG cannot be obtained.

If this step finishes without finding a hole in an expression block, it is guaranteed that an FPG can be obtained because $F_1$ now satisfies all four conditions from Sec.~\ref{ss:FPG} expect for (3), which requires there to be no holes or collisions in the graph. Since no more holes remain, and collisions can always be resolved without producing a hole, (see Table \ref{t:variable node classification}), then it is guaranteed than an FPG can be obtained.

@@

    \item[\bf{Step 2: Collisions}] 
        The second step is to detect collisions and to disconnect precisely the number of connection edges required such that all collisions are resolved without introducting holes. 
        The set of variable nodes at which collisions occur is 
    \begin{equation}
    S_\txt{nodes} = \{v \in V_{M,\txt{var}} \st \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
    \end{equation}
    For each collision node we can construct a set containing the edges directed into the node. The set containing all of these sets is constructed as
    \begin{equation}
        S_\txt{edges} = \big \{ \{(x,y) \in E_M\} \ \big| \ y \in S_\txt{nodes} \big \}
    \end{equation}
    Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. 
    An example collision is shown in Fig.~\ref{f:collision} to exemplify the definition of $S_{\txt{edges},j}$. 
    We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$ (which are sets). 
    We may then construct sets of edges as
    \begin{equation}
    B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } \txt{deg}_u^-(v_j) \leq K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
    \end{equation}
        which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper and lower indegree limits of $v_j$.  
The construction of each $B_j$ corresponds to making a decision about which edges to include and which edges not to include. 
Let the new set of connection edges be
    \begin{equation}
    E_{F_2,\txt{con}} = \{ e \in E_{F_1,\txt{con}} \st e \in B_j \txt{ for some } j \in J\}.
    \end{equation}
The set of all edges is created by removing the connection edges not in $E_{F_2,\txt{con}}$:
    \begin{equation}
    E_{F,2} = E_{F_0} \setminus (E_{M,\txt{con}} \setminus E_{F_2,\txt{con}}),
    \end{equation}
    which gives
    \begin{equation}
    F_2 = (V_{F_0},E_{F_2}).
    \end{equation}
    \begin{figure}[htb!]
        \begin{center}
        \includegraphics[width=3in]{images/analysis_block_collision}
        \end{center}
        \vspace{-20pt}
    \caption{Example variable node indicating a collision.}
    \label{f:collision}
    \end{figure}

    \item[\bf{Step 3: Finalize}] 
        The third and final step is to prune the graph to exclude any analysis blocks that became unneeded after the collisions were resolved.
        This is accomplished by first creating the reverse graph of $F_2$, $R = (V_R,E_R)$, where
	\begin{equation}
	E_R = \{ (x,y) \st (y,x) \in E_{F_2}\}, \ V_R = V_{F_2}
	\end{equation}
	Next, a new node $b$ is added to $V_R$ and edges directed from this node to each of the expression blocks are added:
    \begin{equation}
        \forall i \in I_E,\forall v \in V_{E_i}, \txt{ if } \txt{deg}^+(v)=0 \txt{ then } (b,v) \in E_R
    \end{equation}
        The set of nodes that may be reached from $b$ is then constructed as
    \begin{equation}
        U = \{ v \in V_R \st \exists P \txt{ a path from $b$ to $v$}, \ P \subset R \}.
        \end{equation}
	Because $b$ is directed into only the expression blocks, any path from $b$ necessarily provides a path from at least one expression block, which means the node is being used in the problem formulation.

@@

The fundamental problem graph is then
\begin{equation}
F = (V_F,E_F).
\end{equation}
        If desired, the indices of the analysis blocks implemented in the FPG can then be found as
    \begin{equation}
        I_F = \{ i \in I_A \st G_{A_i} \subset F \}.
    \end{equation}
        Also, the set of connection edges can be extracted by considering only edges whos endpoints are not in the same analysis block:
        \begin{equation}
        E_{F,\txt{con}} = \{(x,y) \in E_{F} \st \sim(x \in V_{A_i} \txt{ and } y \in V_{A_i} \txt{ for some } i \in I_F)\}.
        \end{equation}
\end{description}

This algorithm will always provide an FPG if one exists. If an FPG does not exist, the limiting factors preventing a valid formulation are revealed.

\subsection{Suggested Process for Using the FPG Algorithm}
\label{ss:process}
Section \ref{ss:obtaining FPG} provided an algorithm for obtaining an FPG from the given MCG. However, if an FPG cannot be obtained, this algorithm can be applied as part of a larger process in which the designer changes the problem or the supplied analyses to attempt to obtain an FPG. The following steps detail the suggested procedure for obtaining an FPG:
\begin{enumerate}
\item[\bf{(A)}] Begin with a set of inputs, analysis codes, objectives, and constraints.
\item[\bf{(B)}] Build the MCG as described in Sec.~\ref{ss:MCG}.
\item[\bf{(C)}] Set indegree limits for variable nodes representing local inputs as described in Sec.~\ref{s:indegree-outdegree} and Table \ref{t:variable node classification}
\item[\bf{(D)}] Run the FPG algorithm described in Sec.~\ref{ss:obtaining FPG}. If a valid FPG is unattainable:
		\begin{enumerate}
\item Change the MCG by adding analysis blocks and/or inputs, which means identifying new analyses to include in a potential MDAO workflow.
\item Change indegree limits (see Table \ref{t:variable node classification}).
\end{enumerate}
\item[\bf{(E)}] Repeat from step (A) until an FPG is obtained.
\end{enumerate}
This process illustrates how the algorithm for obtaining a valid FPG is to be applied. However, this process also serves to illustrate how a designer may revist the MCG after an FPG was already produced. For example, if a designer wants to change an input to be multi-fidelity, new analysis blocks can be added to the original MCG and indegree limits can be modified accordingly, which corresponds to steps (A)--(C), and then the algorithm can be applied again.

>>>>

\section{The MCG and FPG}
    \label{s:building graphs}
We now discuss how the graph-based syntax defined in Secs.~\ref{s:syntax definition} and \ref{s:graph representation}  is used to represent an MDAO problem in its entirety. This discussion is based on the MCG and the FPG.

\subsection{Maximal Connectivity Graph}
    \label{ss:MCG}
For a specified MDAO problem, the maximal connectivity graph represents every analysis tool, objective, constraint, and input being considered, as well as every possible interconnection among them.
The definition of the MCG is given via construction.
    To construct the maximal connectivity graph, we presume that a set of analysis tools, inputs, objectives, and constraints are provided. 
    Each of the $m$ analysis codes correspond to an index $i\in I_A$, $I_A=\{1,2,\ldots,m\}$, and are represented by an analysis block graph $G_{A_i}=(V_{A_i},E_{A_i})$.
    Each of the $n$ expression blocks correspond to an index $i\in I_E$, $I_E=\{1,2,\ldots,n\}$, and is represented by expression block graph $G_{E_i}=(V_{E_i},E_{E_i})$.
    Finally, the inputs are represented as a set of variable nodes $V_\txt{in}$.
    We presume that $V_\txt{in}$, each $G_{A_i}$, and each $G_{E_i}$ are given, and that any potential connection between variables is specified in the form of connection edges in the set $E_{M,\txt{con}}$. One method for defining these connections is to use a consistant variable naming convention.

    We may then construct the maximal connectivity graph $M=(V_M,E_M)$ as
    \begin{IEEEeqnarray*}{rCl}
    V_M & = & V_\txt{in} \cup \left( \bigcup_{i \in I_A} V_{A_i} \right) \cup \left( \bigcup_{i \in I_E} V_{E_i} \right), \\
    E_M & = & E_{M,\txt{con}} \cup \left( \bigcup_{i \in I_A} E_{A_i} \right)  \cup \left( \bigcup_{i \in I_E} E_{E_i} \right).
    \end{IEEEeqnarray*}
    The MCG $M$ is uniquely determined by the given set of analysis blocks, inputs, and expression blocks. 
    In the cases where the set of inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.

The nodes and edges in $M$ can be partioned in sets according to their type:
\begin{equation}
V_M = V_{M,\txt{var}} \cup V_{M,\txt{fun}}, \ V_{M,\txt{var}} \cap V_{M,\txt{fun}} = \emptyset,
\end{equation}
where $V_{M,\txt{var}}$ and $V_{M,\txt{mod}}$ are the sets of variable nodes and function nodes, respectively;
\begin{equation}
E_M = E_{M,\txt{con}} \cup E_{M,\txt{fun}}, \ E_{M,\txt{con}} \cap E_{M,\txt{fun}} = \emptyset,
\end{equation}
where $ E_{M,\txt{con}}$ and $E_{M,\txt{fun}}$ are the sets of connection edges and function edges, respectively. These sets will be referenced in the process for obtaining an FPG.

\subsection{Fundamental Problem Graph (FPG)}
    \label{ss:FPG}
    We now define the fundamental problem graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions:
    \begin{enumerate}
    \item[(1)] $F \subset M$ and $G_{E_i} \subset F  \ \forall i \in I_E$
%    \item[(1)] $F \subset M$ and $V_\txt{out} \subset V_F$
    \item[(2)] $\displaystyle{\forall i \in I_A, \txt{ if } F \cap G_{A_i} \neq \emptyset \txt{ then } G_{A_i} \subset F}$
%    \item[(2)] $\displaystyle{\forall i \in I, \txt{ if } F \cap A_i \neq \emptyset \txt{ then } A_i \subset F}$
    \item[(3)] $\displaystyle{\forall v \in V_F \txt{ with }v \in V_{M,\txt{var}}}$ \txt{ and } \\
			 $\displaystyle{ \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F }$ \\
			$\displaystyle{\txt{ from } x \txt{ to } v \txt{ with } x\in V_{E_i}}$
%    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_\txt{out}}$
    \end{enumerate}
    %The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). 
    %Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
    %Requirement (3) stipulates that only the inputs that are being used should be included. 
    %Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
    Condition (1) asserts that only the nodes and edges provided by the maximal connectivity graph can be used in the fundamental problem graph and that every expression block must be included. 
    Condition (2) requires that analysis blocks must be included or excluded in their entirety. 
    Condition (3) stipulates that the number of edges directed into each variable node must be within the lower and upper in-degree limits; 
    if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is the location of a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is the location of a \emph{collision}.
    Lastly, condition (4) ensures that only the nodes that are used in the MDAO problem formulation are included in the FPG by requiring that for every node a reverse path exists from at least one expression block to that node. 
    The reverse graph $R_F$ is obtained from $F$ by simply switching the orientation of every edge in $E_F$.

@@

\subsection{Algorithm for Obtaining the Fundamental Problem Graph from the MCG}
    \label{ss:obtaining FPG}
    In general, there may be zero, one, or many graphs that satisfy the FPG conditions in Sec.~\ref{ss:FPG}.
    Here, we describe an algorithm for obtaining an FPG from a given MCG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. 
    With this approach, the problem is reduced to deciding which connection edges to remove.

@@

        To begin the process, an initial graph is created as a copy of the MCG
        \begin{equation}
        F_0 = M,
        \end{equation}
        where $F_0 = (V_{F_0},E_{F_0})$, and $C_{F_0,\txt{con}}$ is the set of connection edges. 
        The set of variable nodes that are holes is identified as
        \begin{equation}
        H = \{v \in V_{F_0} \st v \in V_{M,\txt{var}} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
        \end{equation}
        which is the set of variable nodes with fewer incoming edges than are allowed by the lower indegree limit.
        An updated set of edges is then created by removing the connection edges preceding or succeeding the analysis block:
\begin{IEEEeqnarray*}{rCl}
        E_{F_1} = E_{F_0} \setminus &\{& (x,y) \in E_{F_0,\txt{con}} \st x \in V_{A_i} \\
& &  \txt{ or } y \in V_{A_i}, \txt{ and } H\cap V_{A_i} \neq \emptyset \},
\end{IEEEeqnarray*}
and the graph is updated as
\begin{equation}
F_1 = (V_{F_0},E_{F_1}).
\end{equation}

Because removing these edges can create new holes, this step must be repeated until no additional holes are found.
        If the hole identification step identifies a variable node in an expression block as hole, meaning $V_{E_i} \cap H \neq \emptyset$ for some $i \in I_E$, then the algorithm terminates because an FPG cannot be obtained.

If this step terminates without identifying a hole in an expression block, it is guaranteed that an FPG can be obtained because $F_1$ now satisfies all conditions from Sec.~\ref{ss:FPG} except for (3), which requires that no holes or collisions be present in the graph. Since the graph is free of holes, and collisions can always be resolved without producing a hole, (see Table \ref{t:variable node classification}), then it is guaranteed than an FPG can be obtained.

@@

    \item[\bf{Step 2: Collisions}] 
        The second step is to detect collisions and to disconnect precisely the number of connection edges required such that all collisions are resolved without introducting holes. 
        The set of variable nodes at which collisions occur is 
    \begin{equation}
    S_\txt{nodes} = \{v \in V_{M,\txt{var}} \st \txt{deg}^-(v) > \txt{deg}_u^-(v) \},
    \end{equation}
where the degree is now calculated with respect to $F_1$.
    For each collision node we can construct a set containing the edges directed into the node. The set containing all of these sets is constructed as
    \begin{equation}
        S_\txt{edges} = \big \{ \{(x,y) \in E_M\} \ \big| \ y \in S_\txt{nodes} \big \}
    \end{equation}
    Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. 
    An example collision is shown in Fig.~\ref{f:collision} to exemplify the definition of $S_{\txt{edges},j}$. 
    $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$ (which are sets). 
    We may then construct sets of edges as
\begin{IEEEeqnarray*}{rCl}
    B_j =  &\big\{& e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \\
& & \txt{ with } \txt{deg}_u^-(v_j) \leq K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
\end{IEEEeqnarray*}

        which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper and lower indegree limits of $v_j$.  
The construction of each $B_j$ corresponds to making a decision about which edges to include and which edges not to include. 
Let the new set of connection edges be denoted
    \begin{equation}
    E_{F_2,\txt{con}} = \{ e \in E_{F_1,\txt{con}} \st e \in B_j \txt{ for some } j \in J\}.
    \end{equation}
The set of all edges is created by removing the connection edges not in $E_{F_2,\txt{con}}$:
    \begin{equation}
    E_{F,2} = E_{F_0} \setminus (E_{M,\txt{con}} \setminus E_{F_2,\txt{con}}),
    \end{equation}
    which gives
    \begin{equation}
    F_2 = (V_{F_0},E_{F_2}).
    \end{equation}
    \begin{figure}[htb!]
        \begin{center}
        \includegraphics[width=3in]{images/analysis_block_collision}
        \end{center}
        \vspace{-20pt}
    \caption{Example variable node indicating a collision.}
    \label{f:collision}
    \end{figure}

    \item[\bf{Step 3: Finalize}] 
        The third and final step is to prune the graph to exclude any analysis blocks that became unneeded after the collisions were resolved.
        This is accomplished by first creating the reverse graph of $F_2$, $R = (V_R,E_R)$, where
	\begin{equation}
	E_R = \{ (x,y) \st (y,x) \in E_{F_2}\}, \ V_R = V_{F_2}
	\end{equation}
	Next, a new node $b$ is added to $V_R$ and edges directed from this node to each of the expression blocks are added:
    \begin{equation}
        \forall i \in I_E,\forall v \in V_{E_i}, \txt{ if } \txt{deg}^-(v)=0 \txt{ then } (b,v) \in E_R.
    \end{equation}
        The set of nodes that may be reached from $b$ is then constructed as
    \begin{equation}
        U = \{ v \in V_R \st \exists P \txt{ a path from $b$ to $v$}, \ P \subset R \}.
        \end{equation}
	Because $b$ is directed into only the expression blocks, any path from $b$ necessarily provides a path from at least one expression block, which means the node is being used in the problem formulation.

@@

The fundamental problem graph is then
\begin{equation}
F = (V_F,E_F).
\end{equation}
        If desired, the set of connection edges can be extracted by considering only edges whose endpoints are not in the same analysis block:
\begin{IEEEeqnarray*}{rCl}
E_{F,\txt{con}} = &\{&(x,y) \in E_{F} \st \sim(x \in V_{A_i} \txt{ and } y \in V_{A_i}  \\
& & \txt{for some } i \in I_F)\}.
\end{IEEEeqnarray*}
\end{description}

This algorithm will always provide an FPG if one exists. If an FPG does not exist, the implementation reveals the limiting factors that prevent a valid FPG from being achieved.

\subsection{Suggested Process for Appyling the FPG Algorithm}
\label{ss:process}
Section \ref{ss:obtaining FPG} provided an algorithm for obtaining an FPG from the given MCG. 
The algorithm can be applied as part of a broader process in which the designer changes the problem formulation or the supplied analysis tools to attempt to obtain an FPG. The following steps detail the suggested procedure for obtaining an FPG:
\begin{enumerate}
\item[\bf{(A)}] Begin with a set of inputs, analysis tools, objectives, and constraints.
\item[\bf{(B)}] Build the MCG as described in Sec.~\ref{ss:MCG}.
\item[\bf{(C)}] Set indegree limits for variable nodes representing local inputs as described in Sec.~\ref{s:indegree-outdegree} and Table \ref{t:variable node classification}.
\item[\bf{(D)}] Run the FPG algorithm described in Sec.~\ref{ss:obtaining FPG}. If a valid FPG is unattainable:
		\begin{enumerate}
\item Change the MCG by adding analysis blocks and/or inputs, which means identifying new analysis tools to include in a potential MDAO workflow.
\item Change indegree limits (see Table \ref{t:variable node classification}).
\end{enumerate}
\item[\bf{(E)}] Repeat from step (A) until an FPG is obtained.
\end{enumerate}
This process illustrates how the algorithm for obtaining a valid FPG can be applied. However, the process also serves to illustrate how a designer may revist the MCG after an FPG has already been produced. For example, if a designer desires to change an input to be multi-fidelity, new analysis blocks can be added to the original MCG, the indegree limits can be modified accordingly, and the algorithm can be applied again.

### 2013-08-07 problem_formulation.tex
files for submitted version and reviewer 3's comments

%\documentclass[referee]{svjour3} % insert '[draft]' option to show overfull boxes
\documentclass[onecolumn]{svjour3} % insert '[draft]' option to show overfull boxes

 \title{A Graph Theory Based Approach to MDAO Problem Formulation}

\author{
  David J. Pate, %
  Justin Gray, %
  Brian J. German 
 }

 \institute { Justin Gray 
	\at NASA Glenn Research Center, Mail Stop 5-11, 21000 
		Brookpark Rd Clevland OH 44135 %
	\and David J. Pate 
	\at Graduate Research Assistant, Georgia Institute of Technology, 270 Ferst Drive, Atlanta, GA, 30332, U.S.A.
	\and Brian German 
	\at Assistant Professor, Georgia Institute of Technology, 270 Ferst Drive, Atlanta, GA, 30332, U.S.A.}

@@

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{caption} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{lscape}
\usepackage{hyperref}
\usepackage{appendix}
\usepackage[section]{placeins}
\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{subfigure}
\usepackage{comment}
\usepackage{booktabs}

@@

\section*{Nomenclature}

\begin{tabular}{l l} 
    AAO      & All-At- \\
    MDAO     & Multidisciplinary Design Analysis and Optimization \\
    FPF      & Fundamental Problem Formulation \\
\end{tabular}

>>>>

%\documentclass[referee]{svjour3} % insert '[draft]' option to show overfull boxes
\documentclass[twocolumn]{svjour3} % insert '[draft]' option to show overfull boxes

 \title{A Graph Theoretic Approach to Problem Formulation for Multidisciplinary Design Analysis and Optimization}

\author{
  David J. Pate,\\ %
  Justin Gray, and\\%
  Brian J. German 
 }

 \institute { David J. Pate 
	\at Graduate Research Assistant, Georgia Institute of Technology, 270 Ferst Drive, Atlanta, GA, 30332, U.S.A.
	\and Justin Gray 
	\at NASA Glenn Research Center, Mail Stop 5-11, 21000 
		Brookpark Rd Clevland OH 44135 %
	\and Brian German 
	\at Assistant Professor, Georgia Institute of Technology, 270 Ferst Drive, Atlanta, GA, 30332, U.S.A.}

@@

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{caption} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{lscape}
\usepackage{appendix}
\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{subfigure}
\usepackage{comment}
\usepackage{booktabs}
\usepackage{multicol}

@@

%\section*{Nomenclature}
%
%\begin{tabular}{l l} 
%    DSM	   & Design structure matrix \\
%    FDT      & Functional dependency table \\
%    FPG      & Fundamental problem graph \\
%    FPF      & Fundamental problem formulation \\
%    MCG      & Maximal connectivity graph \\
%    MDAO     & Multidisciplinary Design Analysis and \\ & Optimization \\
%    MDO      & Multidisciplinary Design Optimization \\
%    REMS     & Reconfigurable multidisciplinary synthesis
%\end{tabular}
